Explaining NonLinear Classification Decisions with Deep Taylor Decomposition by Montavon et al.

### 2022-06-26 Präsentation/Präsentation_23.06.2020.tex
finish presentation and paper

\begin{frame}
\frametitle[Explainability]{Explainability}
\vspace{0.25cm}
\begin{figure}
\includegraphics[width=0.9\textwidth]{image/introduction_image}
\end{figure}
\vspace{0.25cm}
Deep neural networks perform great on a variety of problems \nocite{*}\\
\textbf{but} how can we justify decisions made by complex deep architectures?\\
\end{frame}

@@

\section[Introduction]{Introduction}
\begin{frame}
\frametitle{Introduction} 
\textbf{Deep neural networks revolutionized amongst others the field of}\\
\vspace{0.25cm}
       \begin{columns}[T]
          \column{0.65\textwidth}
             \centering
             \includegraphics[height=5cm, width=8cm]{image/NLP_chatbot}
           \column{0.35\textwidth}
			\begin{itemize}
			\item[--] Image recognition
			\item[--] Natural language processing
			\item[--] Human action recognition
			\item[--] Physics
			\item[--] Finance
			\item[--] ...
			\end{itemize}
         \end{columns} 
\vspace{0.5cm}
With one major drawback $\rightarrow$ \textbf{lack of transparency}
\end{frame}

@@

\end{itemize}

@@

\subsection[Methodology]{Methodology}

@@

\begin{frame}
\frametitle{Sensitivity Analysis}
\vspace{0.25cm}
Root point at infinitesimally small distance from the actual point, i.e. $\xi = x- \delta \frac{\partial f}{\partial x}$, where $\delta$ is small\\
\pause
\vspace{0.1cm}
If we assume a locally constant function we get
\begin{flalign*}
 f(x) & =f(\xi) + \left( \frac{\partial f}{\partial x}\Big|_{x=\xi}\right)^T \cdot (x- (x- \delta \frac{\partial f}{\partial x})) + 0\\
       & = f(\xi) + \delta \left( \frac{\partial f}{\partial x}\right)^T \cdot  \frac{\partial f}{\partial x} + 0\\
       & = f(\xi) + \sum_p \underbrace{ \delta \left( \frac{\partial f}{\partial x}\right)^2}_{R_p} +\ 0
\end{flalign*}

@@

\begin{frame}
\frametitle{Derivation of Propagation Rules}
\vspace{0.3cm}
Given $R_j = \max(0, \sum_i x_i w_{ij} + b_j)$ and $b_j \leq 0$ and a search direction $\{v_i\}^{(j)}$ in the input space such that

@@

\vspace{0.2cm}
If the data point itself is not a root point, i.e. $\sum_i x_i w_{ij} + b_j>0$ the nearest root along $\{v_i\}^{(j)}$ is given by the intersection of equation \eqref{equ:search_direction} and $\sum_i\tilde{x}_i^{(j)} w_{ij} + b_j=0$ which can be resolved to 
\vspace{-0.1cm}
\begin{align*}
0 &= \sum_i x_i w_{ij} + b_j + \sum_i v_i^{(j)} t w_{ij} \\
x_i - \tilde{x}_i^{(j)} & = \frac{\sum_i x_i w_{ij} + b_j}{\sum_i v_i^{(j)} w_{ij}} v_i^{(j)}
\end{align*}

@@

\begin{enumerate}
	\item Search direction $\{v_i\}^{(j)} = w_{ij}$ (gradient of $R_j$)
	\item No domain restriction and for $\tilde{x}_i^{(j)} = x_i - \frac{R_j}{\sum_{i'}w_{i'j}^2}$
	\begin{align*}
		R_j(\tilde{x}) &= \max(0, \sum_i (x_i -\frac{R_j}{\sum_{i'}w_{i'j}^2}) w_{ij}  + b_j) \\
		&= \max(0, \underbrace{\sum_i (x_i w_{ij} + b_j)}_{=R_j} - R_j \underbrace{\frac{\sum_{i}w_{ij}^2}{\sum_{i'}w_{i'j}^2}}_{=1} )= 0 
	\end{align*}
	\vspace{-0.1cm}
	\item Inject search direction in equation \eqref{equ:relevance_model}
\end{enumerate}

@@

\begin{frame}
\frametitle{$z^+$-rule $\mathcal{X}=\mathbb{R}_+^d$}
\vspace{0.5cm}
Search for a root point on the segment $(\{x_i 1_{w_{ij}<0}\},\{x_i\})\subset \mathbb{R}_+^d$\\
\begin{enumerate}
	\item Search direction $\{v_i\}^{(j)} = x_i - x_i 1_{w_{ij}<0} = x_i 1_{w_{ij}\geq0}$
	\item If $\{x_i\} \in \mathbb{R}_+^d$ so is the whole domain, further for $w_{ij}^-=\min(0,w_{ij})$
	\
	\begin{align*}
		R_j(\{x_i1_{w_{ij}<0}\}) =& \max(0, \sum_i x_i 1_{w_{ij}<0} w_{ij} + b_j)\\
		=& \max(0, \underbrace{\sum_i x_i  w_{ij}^-}_{\leq 0} +\ b_j) = 0
	\end{align*}
	\vspace{-0.25cm}
	\item Inject search direction in equation \eqref{equ:relevance_model}
\end{enumerate}

\pause
\begin{block}{$z^+$-rule}
\begin{equation*}
R_i =  \sum_j \frac{x_i w_{ij}^+}{\sum_{i'} x_{i'} w_{i'j}^+} R_j
\end{equation*}
\end{block}

@@

\begin{frame}
\frametitle{$z^b$-rule $\mathcal{X}=\mathcal{B}$}
\vspace{0.4cm}
Given a bounded input space $\mathcal{B} = \{\{x_i\} : \forall_{i=1}^d l_i\leq x_i \leq h_i\} $, with $l_i \leq 0$ and $ h_i \geq 0$ and we chose the segment $(\{l_i 1_{w_{ij}>0} + h_i 1_{w_{ij}<0}\},\{x_i\})\subset \mathcal{B}$\\
\begin{enumerate}
	\item Search direction $\{v_i\}^{(j)} = x_i - l_i 1_{w_{ij}>0} - h_i 1_{w_{ij}<0}$
	\item If $\{x_i\} \in \mathcal{B}$ so is the whole domain and for $w_{ij}^-=\min(0,w_{ij})$, $w_{ij}^+=\max(0,w_{ij})$
	\vspace{-0.25cm}
	\begin{align*}
		R_j(\{l_i 1_{w_{ij}>0} + h_i 1_{w_{ij}<0}\}) &= \max(0, \sum_i l_i 1_{w_{ij}>0} w_{ij} + h_i 1_{w_{ij}<0} w_{ij} + b_j)\\
		&=\max(0, \sum_i \underbrace{l_i w_{ij}^+}_{\leq 0} + \underbrace{h_i w_{ij}^-}_{\leq 0} + b_j)= 0
	\end{align*}
	\vspace{-0.75cm}
	\item Inject search direction in equation \eqref{equ:relevance_model}
\end{enumerate}

@@

 \begin{columns}
          \begin{column}{5cm}
			\begin{align*}
			R_k &= \sum_l \frac{x_k w_{kl}^+}{\sum_{k'} x_{k'} w_{k'l}^+} R_l\\
				&= (\sum_j x_j) \cdot \frac{\norm{\{x_j\}}_p}{\norm{\{x_j\}}_1}   \sum_l \frac{w_{kl}^+R_l}{\sum_{k'} x_{k'} w_{k'l}^+} 
			\end{align*}
            \end{column} 
            \begin{column}{5cm}
			\begin{figure}
			\label{fig1}
			\includegraphics[height=3cm, width=4cm]{image/training-free_relevance_model_3}
			\end{figure}
	\end{column}
\end{columns}

@@

\item Detection layer
\begin{equation*}
R_i = \sum_j \frac{q_{ij}}{\sum_{i'} q_{i'j}} R_j
\end{equation*}
where $q_{ij}= v_{ij}^2$, $q_{ij}= x_i v_{ij}^+$ or $q_{ij}= x_i v_{ij} - l_i v_{ij}^+ - h_i v_{ij}^-$ for the $w^2$-rule, $z^+$-rule and the $z^b$- rule respectively
\end{itemize}
\vspace{0.1cm}
$\rightarrow$ The training-free relevance model is consistent\\

>>>>

\begin{frame}
\frametitle[Explainability]{Explainability}
\vspace{0.25cm}
\begin{figure}
\includegraphics[width=0.9\textwidth]{image/introduction_image}
\end{figure}
\vspace{0.25cm}
Deep neural networks perform great on a variety of problems \nocite{*}\\
\textbf{but} how can we explain decisions made by complex deep architectures?\\
\end{frame}

@@

\section[Introduction]{Introduction}
\begin{frame}
\frametitle{Introduction} 
\textbf{Deep neural networks revolutionized amongst others the field of}\\
\vspace{0.25cm}
       \begin{columns}[T]
          \column{0.62\textwidth}
             \centering
             \includegraphics[height=5cm, width=8cm]{image/NLP_chatbot}
           \column{0.35\textwidth}
			\begin{itemize}
			\item[--] Image recognition
			\item[--] Natural language processing
			\item[--] Human action recognition
			\item[--] Physics
			\item[--] Finance
			\item[--] ...
			\end{itemize}
         \end{columns} 
\pause
\vspace{0.5cm}
With one major drawback $\rightarrow$ \textbf{lack of transparency}
\end{frame}

@@

\pause
\end{itemize}

@@

\subsection[Taylor Expansion and Sensitivity Analysis]{Taylor Expansion and Sensitivity Analysis}

@@

\begin{frame}
\frametitle{Sensitivity Analysis}
\vspace{0.35cm}
Choose a point at infinitesimally small distance from the actual point, i.e. $\xi = x- \delta \frac{\partial f}{\partial x}$, where $\delta$ is small\\
\pause
\vspace{0.35cm}
If we assume a locally constant function we get
\begin{flalign*}
 f(x) & =f(\xi) + \left( \frac{\partial f}{\partial x}\Big|_{x=\xi}\right)^T \cdot (x- (x- \delta \frac{\partial f}{\partial x})) + 0\\
       & = f(\xi) + \delta \left( \frac{\partial f}{\partial x}\right)^T \cdot  \frac{\partial f}{\partial x} + 0\\
       & = f(\xi) + \sum_p \underbrace{ \delta \left( \frac{\partial f}{\partial x}\right)^2}_{R_p} +\ 0
\end{flalign*}

@@

\begin{frame}
\frametitle{Derivation of Propagation Rules}
\vspace{0.5cm}
Given $R_j = \max(0, \sum_i x_i w_{ij} + b_j)$ and $b_j \leq 0$ and a search direction $\{v_i\}^{(j)}$

@@

\pause
\vspace{0.1cm}
If $\sum_i x_i w_{ij} + b_j>0$ the nearest root along the search direction $\{v_i\}^{(j)}$ is given by the intersection of equation \eqref{equ:search_direction} and $\sum_i\tilde{x}_i^{(j)} w_{ij} + b_j=0$  
\begin{align*}
&               & 			0 &= \sum_i x_i w_{ij} + b_j + \sum_i v_i^{(j)} t w_{ij} \\[0.25em]
&\Leftrightarrow& -t & = \frac{\sum_i x_i w_{ij} + b_j}{\sum_i v_i^{(j)} w_{ij}}\\[0.5em]
&\Leftrightarrow& x_i - \tilde{x}_i^{(j)} & = \frac{\sum_i x_i w_{ij} + b_j}{\sum_i v_i^{(j)} w_{ij}} v_i^{(j)}
\end{align*}

@@

\begin{enumerate}
	\item Search direction $\{v_i\}^{(j)} = w_{ij}$ (gradient of $R_j$)
	\item No domain restriction and for $\tilde{x}_i^{(j)} = x_i - \frac{R_j(x_i)}{\sum_{i'}w_{i'j}^2} w_{ij}$
	\begin{align*}
		R_j(\tilde{x}) &= \max(0, \sum_i (x_i -\frac{R_j(x_i) w_{ij}}{\sum_{i'}w_{i'j}^2}) w_{ij}  + b_j) \\
		&= \max(0, \underbrace{\sum_i (x_i w_{ij} + b_j)}_{=R_j(x_i)} - R_j (x_i) \underbrace{\frac{\sum_{i}w_{ij}^2}{\sum_{i'}w_{i'j}^2}}_{=1} )= 0 
	\end{align*}
	\vspace{-0.1cm}
	\item Inject search direction in equation \eqref{equ:relevance_model}
\end{enumerate}

@@

\begin{frame}
\frametitle{$z^+$-rule $\mathcal{X}=\mathbb{R}_+^d$}
\vspace{0.35cm}
Search for a root point on the segment $(\{x_i 1_{w_{ij}<0}\},\{x_i\})\subset \mathbb{R}_+^d$\\
\begin{enumerate}
	\item Search direction $\{v_i\}^{(j)} = x_i - x_i 1_{w_{ij}<0} = x_i 1_{w_{ij}\geq0}$
	\item If $\{x_i\} \in \mathbb{R}_+^d$ so is the whole domain, further for $w_{ij}^-=\min(0,w_{ij})$ and $w_{ij}^+ = \max(0, w_{ij})$
	\vspace{-0.25cm}
	\begin{align*}
		R_j(\{x_i1_{w_{ij}<0}\}) =& \max(0, \sum_i x_i 1_{w_{ij}<0} w_{ij} + b_j)\\
		=& \max(0, \underbrace{\sum_i x_i  w_{ij}^-}_{\leq 0} +\ b_j) = 0
	\end{align*}
	\vspace{-0.25cm}
	\item Inject search direction in equation \eqref{equ:relevance_model}
\end{enumerate}
\vspace{-0.2cm}
\pause
\begin{block}{$z^+$-rule}
\begin{equation*}
R_i =  \sum_j \frac{x_i w_{ij}^+}{\sum_{i'} x_{i'} w_{i'j}^+} R_j
\end{equation*}
\end{block}

@@

\begin{frame}
\frametitle{$z^b$-rule $\mathcal{X}=\mathcal{B}$}
\vspace{0.4cm}
Given a bounded input space $\mathcal{B} = \{\{x_i\} : \forall_{i=1}^d l_i\leq x_i \leq h_i\} $, with $l_i \leq 0$ and $ h_i \geq 0$ and we search on the segment $(\{l_i 1_{w_{ij}>0} + h_i 1_{w_{ij}<0}\},\{x_i\})\subset \mathcal{B}$\\
\begin{enumerate}
	\item Search direction $\{v_i\}^{(j)} = x_i - l_i 1_{w_{ij}>0} - h_i 1_{w_{ij}<0}$
	\item If $\{x_i\} \in \mathcal{B}$ so is the whole domain and for $w_{ij}^-=\min(0,w_{ij})$, $w_{ij}^+=\max(0,w_{ij})$
	\vspace{-0.25cm}
	\begin{align*}
		R_j(\{l_i 1_{w_{ij}>0} + h_i 1_{w_{ij}<0}\}) &= \max(0, \sum_i l_i 1_{w_{ij}>0} w_{ij} + h_i 1_{w_{ij}<0} w_{ij} + b_j)\\
		&=\max(0, \sum_i \underbrace{l_i w_{ij}^+}_{\leq 0} + \underbrace{h_i w_{ij}^-}_{\leq 0} + b_j)= 0
	\end{align*}
	\vspace{-0.75cm}
	\item Inject search direction in equation \eqref{equ:relevance_model}
\end{enumerate}

@@

 \begin{columns}
          \begin{column}{5cm}
			\begin{align*}
			R_k &= \sum_l \frac{x_k w_{kl}^+}{\sum_{k'} x_{k'} w_{k'l}^+} R_l\\
				&= \bigl( \sum_j x_j \bigr)\cdot \frac{\norm{\{x_j\}}_p}{\norm{\{x_j\}}_1}   \sum_l \frac{w_{kl}^+R_l}{\sum_{k'} x_{k'} w_{k'l}^+} 
			\end{align*}
            \end{column} 
            \begin{column}{5cm}
			\begin{figure}
			\label{fig1}
			\includegraphics[height=3cm, width=4cm]{image/training-free_relevance_model_3}
			\end{figure}
	\end{column}
\end{columns}

@@

\item Detection layer
\begin{equation*}
R_i = \sum_j \frac{q_{ij}}{\sum_{i'} q_{i'j}} R_j
\end{equation*}
where $q_{ij}= w_{ij}^2$, $q_{ij}= x_i w_{ij}^+$ or $q_{ij}= x_i w_{ij} - l_i w_{ij}^+ - h_i w_{ij}^-$ for the $w^2$-rule, $z^+$-rule and the $z^b$- rule respectively
\end{itemize}
\vspace{0.1cm}
$\rightarrow$ The training-free relevance model is consistent\\
